---
title: "data_wrangle_woodyplants"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(lubridate)
library(hms)
```

# Read in raw **mammal** data
```{r}
path <- "./woody_plants/"

files <- list.files(path=path, pattern="*.csv")

for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)),
    read_csv(paste(path,file,sep=""), guess_max = 10000))
}
```

# Wrangle temperature data
We always want to keep our **raw** data **raw**, so our next step will be to save a copy of our temperature data data as a different data object. Also for simplicity, we'll call it "temp" (shorter to type!)

First, we'll remove 'NA' values in the tempSingleMean column
```{r}
raw_plants <- vst_apparentindividual %>% 
  filter(!is.na(height))
```

Next, we want to simplify our temperature data to at least a daily value. To do this, we will first use a `mutate()` function to create separate columns for date and time (rather than combined datetime). Next, we `select()` the columns we want to keep (including our new date and time columns)

To know which columns we might want to `select()`, it's helpful to have R re-print the column names for us.
```{r}
names(raw_plants)
```

**Super-important note!** Make sure you include any data quality columns in your initial `select()` function! (in this file, the column is called **finalQF**)

```{r}
plants <- raw_plants %>% 
  mutate(date = as.Date(date),
         month = month(date),
         year = year(date)) %>% 
  select(siteID, year, month, date, height, dataQF)

View(plants)
```

Now, we can group and summarize our data to get (in this example) monthly temperature values
```{r}
monthly_height <- plants  %>% 
  group_by(siteID, year, month) %>% 
  summarize(monthlyMean = mean(height)) %>% 
  arrange(siteID, year, month) %>% 
  mutate(date = make_date(year, month))

View(monthly_height)
```

If you want to save a copy of your *subset* data file, you can run the following command:
```{r}
monthly_height %>% 
  write_csv('./cleaned_data_woodyplants/monthly_height.csv')
```

# Quickly visualize cleaned up temperature data
It's always a good idea to take a quick look at your data to make sure nothing looks *off* (e.g., major outliers, etc)
```{r}
ggplot(monthly_height, aes(x = date,
                     y = monthlyMean, 
                     col = siteID)) +
  geom_point() +
  theme_minimal()
```

