---
title: "Data analysis skeleton: Update to reflect your team's analyses!"
output: html_document
---

# Load necessary packages
```{r}
#install.packages('pacman', repos = "http://cran.us.r-project.org")
pacman::p_load(tidyverse, broom, lubridate, gvlma, car)
```

# Load cleaned data files
## Data type 1: Air temperature
```{r}
temp <- read_csv('./cleaned_data/monthly_temp.csv') %>% 
  rename(monthlyMean_temp = monthlyMean)

glimpse(temp)
```

## Data type 2: Soil moisture
```{r}
soil_moisture <- read_csv('./cleaned_data/monthly_moisture.csv') %>% 
  rename(monthlyMean_soil = monthlyMean)

glimpse(soil_moisture)
```

## Data type 3: Canopy height
```{r}
height <- read_csv('./cleaned_data/monthly_height.csv') %>% 
  rename(monthlyMean_height = monthlyMean)

glimpse(height)
```

## Combine data files based on shared criteria
Type out some details of what the column names are that your data objects have in common. Note that for R to combine them, the column names must be **exactly** the same (e.g., spelling, capitalization, etc.)

```{r}
data <- full_join(temp, soil_moisture) %>%
  full_join(.,height) %>% 
  mutate(siteID = factor(siteID))

print(data)
```

# Statistical analysis:
* If your independent variable is categorical (e.g., siteID) and your dependent variable is continuous (numeric values), you should perform an ANOVA.  
For ANOVA, the R syntax is: **aov(dependent_variable_name ~ independent_variable_name, data = data_object_name)**

* If your independent variable and your dependent variable are both continuous (numeric values), you should perform a linear regression.  
For linear regression, the R syntax is: **lm(dependent_variable_name ~ independent_variable_name, data = data_object_name)**

Note that you should use a different "model_name" for each statistical test you are conducting.

# Summary/descriptive statistics
Often, we report summary statistics like the mean, a measure of variance (standard deviation, standard error, 95% confidence intervals) and the sample size. 

We can do this easily through the use of `dplyr()` commands. In this example, we can calculate average annual values per site using:
```{r}
summary <- data %>%
  mutate(year = year(date)) %>% 
  group_by(siteID, year) %>% 
  summarize(n = n(),  # Sample size
            Mean_temp = mean(monthlyMean_temp, na.rm=T), sd_temp = sd(monthlyMean_temp, na.rm=T), se_temp = sd_temp/sqrt(n), # Temp
            Mean_soil = mean(monthlyMean_soil, na.rm=T), sd_soil = sd(monthlyMean_soil, na.rm=T), se_soil = sd_soil/sqrt(n), # Soil moisture
            Mean_height = mean(monthlyMean_height, na.rm=T), sd_height = sd(monthlyMean_height, na.rm=T), se_height = sd_height/sqrt(n)) # Canopy height

print(summary)
```

To simplify reporting, we can also ask R to round all of these calculated values to a specific number of decimal places:
```{r}
summary <- as_tibble(summary) %>% 
  mutate_at(vars(Mean_temp:se_height), funs(round(.,2))) # Round each of our calculated columns to 2 decimals

print(summary)
```

# ANOVA
Write a brief statement about what you're comparing, ecologically, with each planned analysis.

First, let's make a quick plot of our site-level response variable; after we run our ANOVA we can visually compare this graph our statistical test results:
```{r}
level_order <- c('ONAQ', 'DSNY', 'JERC', 'ABBY','SCBI','HARV','MLBS','UNDE')

ggplot(data = data, aes(x = factor(siteID, level=level_order), # name the variable that goes on the x-axis
                        y = monthlyMean_height,  # name the variable that goes on the y-axis
                        fill = siteID)) + # name the variable to differentiate colors
  geom_boxplot() +  # Graph these variables using boxplots
  geom_point(alpha = 0.5) +
  theme_minimal()
```
```{r}
ggplot(data = data, aes(x = siteID, # name the variable that goes on the x-axis
                        y = monthlyMean_temp,  # name the variable that goes on the y-axis
                        fill = siteID)) + # name the variable to differentiate colors
  geom_boxplot() +  # Graph these variables using boxplots
  geom_point(alpha = 0.5) +
  theme_minimal()
```
```{r}
ggplot(data = data, aes(x = siteID, # name the variable that goes on the x-axis
                        y = monthlyMean_soil,  # name the variable that goes on the y-axis
                        fill = siteID)) + # name the variable to differentiate colors
  geom_boxplot() +  # Graph these variables using boxplots
  geom_point(alpha = 0.5) +
  theme_minimal()
```

Now, we're ready to define our ANOVA model:
```{r}
height_model <- aov(monthlyMean_height ~ siteID, data = data)
```

```{r}
temp_model <- aov(monthlyMean_temp ~ siteID, data = data)
```

```{r}
soil_model <- aov(monthlyMean_soil ~ siteID, data = data)
```

Next, we can use this quick command to check the ANOVA model assumptions:
```{r, warning=F}
# Check for homogeneity of variance: 
leveneTest(monthlyMean_height ~ siteID, data = data)
# If the p-value of Levene's test is greater than 0.05, we have not violated the assumption of homogeneity of variance

# Check for normality of residuals:
## 1) Extract the model residuals
aov_residuals <- residuals(object = height_model)

## 2) Run Shapiro-Wilk test
shapiro.test(x = aov_residuals )

# Visualize model checks as graphs
#check_model(height_model)
```
```{r, warning=F}
# Check for homogeneity of variance: 
leveneTest(monthlyMean_temp ~ siteID, data = data)
# If the p-value of Levene's test is greater than 0.05, we have not violated the assumption of homogeneity of variance
```


```{r, warning=F}
# Check for normality of residuals:
## 1) Extract the model residuals
aov_residuals <- residuals(object = height_model)

## 2) Run Shapiro-Wilk test
shapiro.test(x = aov_residuals )

# Visualize model checks as graphs
#check_model(temp_model)
```

*(If our data violate the homogeneity of variance assumption, we can instead do a Welch one-way test using the `oneway.test()` function below)*
```{r, eval=F}
oneway.test(monthlyMean_height ~ siteID, data = data, var.equal = FALSE)

pairwise.t.test(data$monthlyMean_height,
                data$siteID,
                 p.adjust.method = "BH", pool.sd = FALSE)
```

If our model assumptions seem to be satisfied, we can look at the ANOVA output! 
```{r}
summary(height_model)
```
If the model p-value is smaller than 0.05 (our standard alpha), we reject the null hypothesis and conclude that there is a significant difference in our dependent variable between *at least two* sites.

But which site(s) are different? We'll use a post-hoc Tukey test to find out!
```{r}
tukey <- tidy(TukeyHSD(height_model))

# Filter for only statistically significant (p <0.05) contrasts
sig <- tukey %>% 
  filter(adj.p.value < 0.05)

print(sig)
```
Note that the "estimate" column tells you how different the mean values of each pair of sites is.

# Correlation &Linear regression
Write a brief statement about what you're comparing, ecologically, with each planned analysis.
We are comparing air temperature and canopy height, as well as soil moisture and canopy height. 

First, let's make a quick plot to see what the data look like
```{r}
ggplot(data = data, aes(x = monthlyMean_temp, # name the variable that goes on the x-axis
                                    y = monthlyMean_height,  # name the variable that goes on the y-axis
                                    col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16) +  # Graph these variables as a scatterplot
  theme_minimal()
```

```{r}
ggplot(data = data, aes(x = monthlyMean_soil, # name the variable that goes on the x-axis
                                    y = monthlyMean_height,  # name the variable that goes on the y-axis
                                    col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16) +  # Graph these variables as a scatterplot
  theme_minimal()
```


We can calculate a correlation coefficient to quantify the strength and direction of association between the two variables.

The types of correlation you can use are:  

* Pearson: Parametric correlation (requires data be normally distributed) for continuous data   

* Spearman: Non-parametric (e.g., if your data are not normally distributed) correlation can also take ordinal data

```{r}
cor_alpha <- cor.test(data$monthlyMean_soil, 
                      data$monthlyMean_height, 
                      method = "pearson") # can also use method = "spearman"

tidy(cor_alpha)
```
```{r}
cor_alpha <- cor.test(data$monthlyMean_temp, 
                      data$monthlyMean_height, 
                      method = "pearson") # can also use method = "spearman"

tidy(cor_alpha)
```

Now, let's set up and run a linear model to conduct a linear regression.
```{r}
soil_model2 <- lm(monthlyMean_height ~ monthlyMean_soil, data = data)
```

```{r}
temp_model2 <- lm(monthlyMean_height ~ monthlyMean_temp, data = data)
```

Next, we can use a few quick commands to check the linear regression model assumptions:
```{r, warning=F}
gvlma(soil_model2)
```

```{r, warning=F}
gvlma(temp_model2)
```

If our model assumptions seem to be satisfied, we can look at the ANOVA output! 
```{r}
summary(soil_model2)
```
```{r}
summary(temp_model2)
```

And make a quick plot with the calculated best fit lines for each site. Remember that ecological data reporting convention is to **omit** the trend line for non-significant relationships.

```{r, message=F}
ggplot(data = data, aes(y = monthlyMean_height, # name the variable that goes on the y-axis
                                    x = monthlyMean_temp,  # name the variable that goes on the x-axis
                                    col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16) +  # Graph these variables as a scatterplot
  geom_smooth(method = 'lm', aes(group=1), col = 'black', lty = 2) +  # lty=1 for solid line
  #geom_smooth(method= 'lm', se=F) +
  theme_minimal()
```

```{r, message=F}
ggplot(data = data, aes(x = monthlyMean_height, # name the variable that goes on the x-axis
                                    y = monthlyMean_soil,  # name the variable that goes on the y-axis
                                    col = siteID)) + # name the variable to differentiate colors
  geom_point(pch = 16) +  # Graph these variables as a scatterplot
  geom_smooth(method = 'lm', aes(group=1), col = 'black', lty = 2) +  # lty=1 for solid line
  #geom_abline(intercept = 6.81098, slope = -0.08871) +  # Alternate method than geom_smooth to specify line slope & intercept
  theme_minimal()
```