---
title: "Tree group Wrangle"
output: html_document
---

#Install & load packages
As always, we begin with the `install_packages()` (if needed) and `library()` commands to install and load individual packages, respectively.
```{r, message = FALSE}
library(tidyverse)
library(lubridate)
library(hms)
```

# Read in raw_soil_moisture data
```{r}
path <- "./raw_soil_moisture/"

files <- list.files(path=path, pattern="*.csv")

for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)),
    read_csv(paste(path,file,sep=""), guess_max = 10000))
}
```

# Wrangle Soil Moisture data
We always want to keep our **raw** data **raw**, so our next step will be to save a copy of our temperature data data as a different data object. Also for simplicity, we'll call it "temp" (shorter to type!)

First, we'll remove 'NA' values in the tempSingleMean column
```{r}
soilmoisture<- SWS_30_minute %>% 
  filter(!is.na(VSWCMean))
```


Next, we want to simplify our temperature data to at least a daily value. To do this, we will first use a `mutate()` function to create separate columns for date and time (rather than combined datetime). Next, we `select()` the columns we want to keep (including our new date and time columns)

To know which columns we might want to `select()`, it's helpful to have R re-print the column names for us.
```{r}
names(soilmoisture)
```

**Super-important note!** Make sure you include any data quality columns in your initial `select()` function! (in this file, the column is called **finalQF**)
```{r}
moisture_sub <- soilmoisture %>% 
  mutate(date = as.Date(startDateTime),
         month = month(date),
         year = year(date),
         time = as_hms(startDateTime)) %>% 
  select(VSWCMean,siteID,date,month,year,VSWCFinalQF)

View(moisture_sub)
```
We should only retain moisture values that do **not** have a failing quality flag (so our filter criterion would be: finalQF != 1 *or* finalQF == 0)
```{r}
moisture_sub <- moisture_sub %>% 
  filter(VSWCFinalQF != 1)
```

Now, we can group and summarize our data to get (in this example) monthly temperature values
```{r}
monthly_moisture <- moisture_sub  %>% 
  group_by(siteID, year, month) %>% 
  summarize(monthlyMean = mean(VSWCMean)) %>% 
  arrange(siteID, year, month) %>% 
  mutate(date = make_date(year, month))

View(monthly_moisture)
```

If you want to save a copy of your *subset* data file, you can run the following command:
```{r}
monthly_moisture %>% 
  write_csv('./cleaned_data_woodyplants/monthly_moisture.csv')
```