---
title: "Data exploration example"
output: html_document
---

# Install & load packages  
As always, we begin with the `install_packages()` (if needed) and `library()` commands to install and load individual packages, respectively.
```{r, message = FALSE}
#install.packages('tidyverse', repos = "http://cran.us.r-project.org")
#install.packages('lubridate', repos = "http://cran.us.r-project.org")
library(tidyverse)
library(lubridate)
```

# Read in your raw NEON data
```{r}
path <- "./data/SCBI_mammals_raw/"

files <- list.files(path=path, pattern="*.csv")
```

Now we'll use the following command to read in all our data files from the folder we specified above! 
```{r, warning = FALSE}
for(file in files){
  perpos <- which(strsplit(file, "")[[1]]==".")
  assign(
    gsub(" ","",substr(file, 1, perpos-1)),
    read_csv(paste(path,file,sep="")))
}
```

# Explore a data file
We always want to keep our **raw** data **raw**, so our next step will be to save a copy of our raw mammal trapping data as a different data object. Also for simplicity, we'll call it "mamdata" (shorter to type!)
```{r}
mamdata <- mam_pertrapnight
```

Now we can use some of the `dplyr` functions we explored last week to begin filtering, organizing, and exploring the data!
```{r}
captures <- mamdata %>% 
  select(siteID, plotID, nlcdClass, collectDate, taxonID, tagID, sex, lifeStage) %>% # Select certain columns
  filter(!is.na(tagID))                                                              # Only keep rows where "tagID" has data
```

If you want to save a copy of your *subset* data file, you can run the following command:
```{r}
captures %>% 
  write_csv('./data/subset_mammal_data.csv')
```

